---
title: "Reproducible Research, Assignment One"
author: "AJ"
date: "October 5, 2016"
output: html_document
---

##Introduction
JHU Coursera -- Reproducible Research (Course 5), Week 2 Assignment

##Prep Enviroment
```{r}
rm(list = ls())
setwd("C:/Users/akash/Desktop/StatsCourses/JHU_Specialization/Course5/w2")
library(ggplot2)
library(knitr)
opts_chunk$set(echo = TRUE, results = 'hold')
```

##Load Data

####Assumes datafile is in setwd
```{r}
rdata <- read.csv('activity.csv', header = TRUE, sep = ",",                   
            colClasses=c("numeric", "character", "numeric"))
```

##Tidy The Data
```{r}
rdata$date <- as.Date(rdata$date, format = "%Y-%m-%d")
rdata$interval <- as.factor(rdata$interval)
```

####Check
```{r}
str(rdata)
```

##**What Is The Mean Total Number Of Steps Taken Per Day?**

####Total Number Of Steps Taken Per Day
```{r}
steps_per_day <- aggregate(steps ~ date, rdata, sum)
colnames(steps_per_day) <- c("date","steps")
head(steps_per_day)
```

####Histogram: Total Number of Steps Taken Per Day
```{r}
ggplot(steps_per_day, aes(x = steps)) + 
    geom_histogram(fill = "blue4", binwidth = 1000) + 
    labs(title="Histogram of Steps Taken per Day",
    x = "Number of Steps per Day",
    y = "Number of times in a day(Count)") + theme_bw() 
```

####Mean And Median: Total Number Of Steps Taken Per Day
```{r}
steps_mean   <- mean(steps_per_day$steps, na.rm=TRUE)
steps_median <- median(steps_per_day$steps, na.rm=TRUE)
```
The mean is **`r format(steps_mean,digits = 8)`** and median is **`r format(steps_median,digits = 8)`**.

##**What is the Average Daily Activity Pattern?**

####Time Series Plot: Average Number Of Steps Taken (Averaged Across All Days) Vs the 5-Minute Intervals
```{r}
steps_per_interval <- aggregate(rdata$steps, 
    by = list(interval = rdata$interval), FUN=mean, na.rm=TRUE)
steps_per_interval$interval <- as.integer(levels(steps_per_interval$interval)[steps_per_interval$interval])
colnames(steps_per_interval) <- c("interval", "steps")
ggplot(steps_per_interval, aes(x=interval, y=steps)) + 
    geom_line(color="purple", size=1) +  
    labs(title="Average Daily Activity Pattern", x="Interval",
    y="Number of steps") + theme_bw()
```

####Max 5 Minute Interva;
```{r}
max_interval <- steps_per_interval[which.max(steps_per_interval$steps),]
```
The **`r max_interval$interval`<sup>th</sup>** interval has maximum **`r round(max_interval$steps)`** steps.

##**Imputing Missing Values**

######Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

####Total Number Of Missing Values
```{r}
missing_vals <- sum(is.na(rdata$steps))
```
The total number of missing values are **`r missing_vals`**.

####Solution: How To Fix The N/As

```{r}
na_fill <- function(data, pervalue) {
        na_index <- which(is.na(data$steps))
        na_replace <- unlist(lapply(na_index, FUN=function(idx){
                interval = data[idx,]$interval
                pervalue[pervalue$interval == interval,]$steps
        }))
        fill_steps <- data$steps
        fill_steps[na_index] <- na_replace
        fill_steps
}

rdata_fill <- data.frame(  
        steps = na_fill(rdata, steps_per_interval),  
        date = rdata$date,  
        interval = rdata$interval)
str(rdata_fill)
```

####Check

```{r}
sum(is.na(rdata_fill$steps))
```
Zero output shows that there are *no missing values*.

####Histogram: Total Number of Steps Taken Per Day (With Filled In N/As)
```{r}
fill_steps_per_day <- aggregate(steps ~ date, rdata_fill, sum)
colnames(fill_steps_per_day) <- c("date","steps")
ggplot(fill_steps_per_day, aes(x = steps)) + 
       geom_histogram(fill = "turquoise4", binwidth = 1000) + 
       labs(title="Histogram of Steps Taken per Day", 
       x = "Number of Steps per Day", 
       y = "Number of times in a day(Count)") + theme_bw() 

```

####Mean And Median: Total Number Of Steps Taken Per Day (With Filled In N/As)
```{r}
steps_mean_fill   <- mean(fill_steps_per_day$steps, na.rm=TRUE)
steps_median_fill <- median(fill_steps_per_day$steps, na.rm=TRUE)
```
The mean is **`r format(steps_mean_fill,digits = 8)`** and median is **`r format(steps_median_fill,digits = 8)`**.

####Do these values differ from the estimates from the first part of the assignment?

#####Before Filling In N/As
    Mean  : `r format(steps_mean,digits = 8)`
    Median: `r format(steps_median,digits = 8)`
    
    
#####After Filling In N/As
    Mean  : `r format(steps_mean_fill,digits = 8)`
    Median: `r format(steps_median_fill,digits = 8)`
The mean is virtually the same in Before vs After but the median slightly increases.  

##**Are There Differences In Activity Patterns Between Weekdays And Weekends?**

```{r}
weekdays_steps <- function(data) {
    weekdays_steps <- aggregate(data$steps, by=list(interval = data$interval),
                          FUN=mean, na.rm=T)
 
    weekdays_steps$interval <- 
            as.integer(levels(weekdays_steps$interval)[weekdays_steps$interval])
    colnames(weekdays_steps) <- c("interval", "steps")
    weekdays_steps
}

data_by_weekdays <- function(data) {
    data$weekday <- 
            as.factor(weekdays(data$date)) # weekdays
    weekend_data <- subset(data, weekday %in% c("Saturday","Sunday"))
    weekday_data <- subset(data, !weekday %in% c("Saturday","Sunday"))
    
    weekend_steps <- weekdays_steps(weekend_data)
    weekday_steps <- weekdays_steps(weekday_data)
    
    weekend_steps$dayofweek <- rep("weekend", nrow(weekend_steps))
    weekday_steps$dayofweek <- rep("weekday", nrow(weekday_steps))
    
    data_by_weekdays <- rbind(weekend_steps, weekday_steps)
    data_by_weekdays$dayofweek <- as.factor(data_by_weekdays$dayofweek)
    data_by_weekdays
}

data_weekdays <- data_by_weekdays(rdata_fill)
```
####Panel Plot: Weekdays Vs Weekends
```{r plot_weekdays}
ggplot(data_weekdays, aes(x=interval, y=steps)) + 
        geom_line(color="darksalmon") + 
        facet_wrap(~ dayofweek, nrow=2, ncol=1) +
        labs(x="Interval", y="Number of steps") +
        theme_bw()

```

#####Weekdays has the highest peak, but Weekend has more steps over 100. 

######The End.






